---
title: "Reproducible document"
author: "Lærke Overgaard"
format: html
---

## Basics of R

```{r}
1 + 1
```
## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
```
## Showing the data

```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
#| fig-cap: "Distrubution of BMI"
#| label: fig-bmi-histo

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
@fig-bmi-histo

## Excercise discrete plots
```{r}
#| fig-cap: "Barchart of Group"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

```{r}
post_meal_data |> # the |> takes data into next step. Same as the approach above but nicer to look at.
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```
```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Put two plots side by side

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

```{r side-by-side figs}
#| label: fig-bmi-group
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap: #In order from left to right
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```
## Styling

```{r}
# R will fix some style issues by using the "style selection" function found in the panel
```

## Plotting two continous variables
```{r}
#| fig-cap: "scatterplot showing BMI against AUC of blood glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```
## Pushing changes to remote

## Pulling changes locally

## Selecting data
```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group) # here we select all data except group

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```
## Renaming specific columns

```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg) # New name to the left

small_post_meal_data |> # renaming multiple characters at once
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 3.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)

post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```
## Filtering data by row
```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")
```
## Logical conditions
```{r}
TRUE & TRUE
TRUE & FALSE
FALSE & FALSE
TRUE & NA
# if you use the "AND" it will only evaluate TRUE if both statements are ture

TRUE | TRUE
TRUE | FALSE
FALSE | FALSE
TRUE | NA
# If you use OR only one condition needs to be true to evaluate as TRUE
```
```{r}
small_post_meal_data |>
  filter(Group == "CTR", BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```

## Modify or add column
```{r}
tidier_post_meal_data |>
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
# Mutate creates new column (in this case ratio between two other columns)
# If else is in the form: condition, what to do with true, what to do with false
```



## About me

- Lærke Karner Overgaard
- Research unit of Clinical Pharmacology
- Department of clinical research
- University of Southern Denmark

I am a **pharmacist** and my research is in *insulin adherence*

## Simple code

```{r}
3 * 3
```

## Testing for Git

It is cold today
