---
title: "Reproducible document"
author: "Lærke Overgaard"
format: html
---

## Basics of R

```{r}
1 + 1
```
## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)
```
## Showing the data

```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
#| fig-cap: "Distrubution of BMI"
#| label: fig-bmi-histo

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
@fig-bmi-histo

## Excercise discrete plots
```{r}
#| fig-cap: "Barchart of Group"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

```{r}
post_meal_data |> # the |> takes data into next step. Same as the approach above but nicer to look at.
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```
```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Put two plots side by side

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

```{r side-by-side figs}
#| label: fig-bmi-group
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap: #In order from left to right
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```
## Styling

```{r}
# R will fix some style issues by using the "style selection" function found in the panel
```

## Plotting two continous variables
```{r}
#| fig-cap: "scatterplot showing BMI against AUC of blood glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() + 
  geom_smooth()
```
## Pushing changes to remote



## About me

- Lærke Karner Overgaard
- Research unit of Clinical Pharmacology
- Department of clinical research
- University of Southern Denmark

I am a **pharmacist** and my research is in *insulin adherence*

## Simple code

```{r}
3 * 3
```

## Testing for Git

It is cold today
